---
title: "Untitled"
author: "Kiera Matiska"
date: "1/31/2022"
output: 
  html_document:
  code_folding: hide
---

# Overview:


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

# attach packages
library(tidyverse)
library(here)
library(GGally)
library(AICcmodavg)
library(broom)
library(caret)
library(jtools)
library(kableExtra)
```

## Read in Data

```{r}
palmetto <- read_csv("palmetto.csv") # read in data
```

## Wrangle and Visualize Data {.tabset .tabset-fade}

Create a new data set with the species name in replace of 1 and 2 in species column

```{r}
palmetto_species <- palmetto %>% # define subset
  mutate(species = case_when( # rename species
    species == "1" ~ "Serenoa repens",
    species == "2" ~ "Sabal etonia"
  ))
```

Use `ggpairs` to compare data

```{r, include = FALSE}
palmetto_species %>% 
  select(species, height:green_lvs) %>% 
  ggpairs(aes(color = species)) # compare variables
```

Create a subset with only:

- species
- height
- canopy length
- canopy width
- green leaves

```{r}
palmetto_subset <- palmetto_species %>% # define new subset
  select(species, height:green_lvs) %>% # only keep these columns
  drop_na() # drop na values
```

### Plot 1

```{r, include = FALSE}
# wrangle data
palmetto_sum1 <- palmetto_subset %>% 
  group_by(species) %>% 
  summarize(max_height = max(height), # find max height
            min_height = min(height), # find min height
            mean_height = mean(height), # calculate mean height
            sd_height = sd(height), # calculate the dtandard deviation
            species_count = n()) # count the number of individuals of each species
```

```{r}
# histogram on height and number of individuals
ggplot(data = palmetto_subset, # define ggplot
       aes(x = height)) + 
  geom_histogram(bins = 20, fill = "darkolivegreen4", color = "darkolivegreen") + # define histogram and change colors
  facet_wrap(~ species) + # separate data by species
  labs(title = "Number of Individuals of Each\nPalmetto Species at Different Heights", # add title
       x = "Height (cm)", # add x-axis label
       y = "Number of Individuals") + # add y-axis label
  theme(plot.title = element_text(hjust = 0.5), # center title
        plot.background = element_rect(fill = "aliceblue")) # change background color
```

**Figure 1.** Number of Individuals of Each Palmetto Species at Different Heights. The population count of *Sabal etonia* (species 1) and *Serenoa repens* (species 2) is `r palmetto_sum1$species_count[1]` and `r palmetto_sum1$species_count[2]`, respectively. Species 1 has a maximum height at `r palmetto_sum1$max_height[1]` cm, and minimum height at `r palmetto_sum1$min_height[1]` cm, and a mean height at `r palmetto_sum1$mean_height[1]` cm. Species 2 has a maximum height at `r palmetto_sum1$max_height[2]` cm, and minimum height at `r palmetto_sum1$min_height[2]` cm, and a mean height at `r palmetto_sum1$mean_height[2]` cm. For both species 1 and species 2, they have a standard deviation of `r palmetto_sum1$sd_height[1]` and `r palmetto_sum1$sd_height[2]`, respectively.

### Plot 2

```{r, include = FALSE}
# wrangle data for plot 2
palmetto_sum2 <- palmetto_subset %>% # define new subset
  group_by(species) %>% 
  summarize(max_leaves = max(green_lvs), # find max number of leaves
            min_leaves = min(green_lvs), # find minimum number of leaves
            median_leaves = median(green_lvs), # calculate median number of leaves
            mean_leaves = mean(green_lvs), # calculate mean number of leaves
            count = n()) # count species
```

```{r}
# create a box plot showing green leave count differences by species
ggplot(data = palmetto_subset, # define ggplot
       aes(x = species, y = green_lvs)) +
  geom_boxplot(color = "black") + # define box plot and chose color
  labs(title = "Number of Green Leaves \nBetween Two Palmetto Species", # add title
       x = "Species", # add x-axis label
       y = "Count of Green Leaves") + # add y-axis label
  theme_gray() + # change theme
  theme(plot.title = element_text(hjust = 0.5), # center title
        plot.background = element_rect(fill = "aliceblue")) # change background color
```

**Figure 2.** Number of Green Leaves Between Two Palmetto Species. The number of individuals of *Sabal etonia* (species 1) and *Seriona repens* (species 2) is `r palmetto_sum2$count[1]` and `r palmetto_sum2$count[1]`, respectively. Species 1 has four outliers and species 2 has 8 outliers. For species 1, the maximum number of leaves an individual has is `r palmetto_sum2$max_leaves[1]`, the minimum number of leaves an individual has is `r palmetto_sum2$min_leaves[1]`, and the median number of leaves for all of the individuals is `r palmetto_sum2$median_leaves[1]`. This first quantile of species 1 is 3 leaves per individual and the third quantile is 5 leaves per individual. For species 2, the maximum number of leaves an individual has is `r palmetto_sum2$max_leaves[2]`, the minimum number of leaves an individual has is `r palmetto_sum2$min_leaves[2]`, and the median number of leaves for all of the individuals is `r palmetto_sum2$median_leaves[2]`. This first quantile of species 2 is 3 leaves per individual and the third quantile is 9 leaves per individual.

### Plot 3

```{r}
# create a scatter plot plotting length against width colored by the number of green leaves an individual has
ggplot(data = palmetto_subset, # define data set
       aes(x = length, y = width)) + 
  geom_point(aes(color = green_lvs), # define scatter plot and change colors
             size = 0.3) + # resize dots
  facet_wrap(~ species) + # separate data by species
  scale_color_gradient(low = "green", high = "yellow") + # change gradient color
  labs(title = "Width of Individual Palmetto Species and\ntheir Green Leave Count Based on Length", # add title
       x = "Length(cm)", # add x-axis
       y = "Width (cm)") + # add y-axis
  theme(plot.title = element_text(hjust = 0.5), # center title
        plot.background = element_rect(fill = "aliceblue")) # change background color
```

**Figure 3.** Width of Individual Palmetto Species and their Green Leave Count Based on Length. Width is measured in centimeters and is a function of the length of an individual, also in centimeters. Number of green leaves is also graphed based on both length and width, with `r palmetto_sum2$max_leaves[2]` being the maximum number of leaves and `r palmetto_sum2$min_leaves[1]` being the smallest number of leaves on an individual.

## Conclusion

Based on these graphs, it looks like length, width, and green leaves will be good predictor variables for the determination of which species the palmetto is.

## Binary Linear Regression Model

Create a binary linear regression model subset

```{r}
palmetto_blr_subset <- palmetto_subset %>% # create a blr subset
  mutate(species = fct_drop(species)) %>% # create a factor of species
  drop_na() # drop na values
```

```{r}
# levels: 1. Sabal etonia, 2. Serenow repens

f1 <- species ~ height + length + width + green_lvs # formula 1
f2 <- species ~ height + width + green_lvs # formula 2

# run generalized linear model (glm) of model 1
palm_species_blr1 <- glm(formula = f1,
                         data = palmetto_blr_subset,
                         family = "binomial")
# run glm of model 2
palm_species_blr2 <- glm(formula = f2, 
                         data = palmetto_blr_subset,
                         family = "binomial")

# palm_species_blr1
# palm_species_blr2

# summary(palm_species_blr1)
# summary(palm_species_blr2)
```

Find AIC values

```{r}
AICc(palm_species_blr1) # 5194.57
AICc(palm_species_blr2) # 5987.48

aictab(list(palm_species_blr1, palm_species_blr2)) # Delta AIC: 792.91
```

Currently, model 1 is the best model not only because it has the lowest AIC value, but the p-values all show significance among each variable. Model 2 has an AIC value that is almost 800 above model 1 and the only two of the three variables show significance.

### Model Selection

Use `caret` package to automate cross validation

```{r}
set.seed(123) # so we don't get random numbers each time we run it

tr_ctrl <- trainControl(method = "repeatedcv", number = 10, repeats = 10) # set train control model

# train the model
model1 <- train(f1, data = palmetto_blr_subset, # train model 1
                method = "glm", family = "binomial",
                trControl = tr_ctrl)
model1 # return model 1

model2 <- train(f2, data = palmetto_blr_subset, # train model 2
                method = "glm", family = "binomial",
                trControl = tr_ctrl)

model2 # return model 2
```

```{r}
# final model: model 1; run glm
final_palmetto_model <- glm(formula = f1, 
                            data = palmetto_blr_subset,
                            family = "binomial")

summary_final <- summary(final_palmetto_model) # create a summary table

# find fitted values using broom::augment
final_fitted <- final_palmetto_model %>% 
  broom::augment(type.predict = "response") %>% 
  mutate(fitted = .fitted) %>% 
  mutate(species = as.character(species))

# create a new table with the odds of the model
final_fitted_odds <- final_fitted %>% # define new subset
  select(species:green_lvs, fitted) %>% # choose certain columns for new subset
  mutate(fitted = fitted * 100) %>% # find percentage
  mutate(suspected_species = case_when(
    fitted >= 50 ~ "Serenoa repens",
    fitted < 50 ~ "Sabal etonia"
  ))
```

```{r}
model_correct <- final_fitted_odds %>% # define new subset
  mutate(species_correct = case_when( # determine whether the model calculated the right or wrong species
    species == "Serenoa repens" & suspected_species == "Serenoa repens" ~ 1,
    species == "Serenoa repens" & suspected_species != "Serenoa repens" ~ 0,
    species == "Sabal etonia" & suspected_species == "Sabal etonia" ~ 4,
    species == "Sabal etonia" & suspected_species != "Sabal etonia" ~ 3,
  ))

# create a summary table with the right and wrong species
model_sum <- model_correct %>% 
  group_by(species_correct) %>% 
  summarize(count = n())

model_sum_sub <- model_sum %>% 
  mutate(species_correct = case_when(
    species_correct == "1" ~ "Serenoa repens True",
    species_correct == "0" ~ "Serenoa repens False",
    species_correct == "4" ~ "Sabal etonia True",
    species_correct == "3" ~ "Sabal etonia False"
  )) %>% 
  mutate(species_percent = case_when(
    count == "564" ~ round(count[1]/(count[1] + count[2]) * 100, 2),
    count == "5548" ~ round(count[2]/(count[1] + count[2]) * 100, 2),
    count == "454" ~ round(count[3]/(count[3] + count[4]) * 100, 2),
    count == "5701" ~ round(count[4]/(count[3] + count[4]) * 100, 2)
  ))

model_sum_sub %>% 
  kbl(caption = "Percentages of the Amount of Correct and Incorrect Guesses Made by the Model.",
      col.names = "Species Status",
      "Count",
      "Percentage (%)") %>% 
  kable_styling()
```

```{r}
final_palm_tidy <- broom::tidy(final_palmetto_model) %>% 
  mutate(term = case_when(
    term == "(Intercept)" ~ "Intercept",
    term == "height" ~ "Height",
    term == "length" ~ "Length",
    term == "width" ~ "Width",
    term == "green_lvs" ~ "Green Leaves"
  ))

final_palm_tidy %>% 
  kable(caption = "Table 1: Binary Linear Regression Results for Model 1",
      col.names = "Coefficient",
      "Estimate",
      "Standard Error",
      "Statistic",
      "P Value",
      digits = c(0, 2, 3, 4))
```

## Conclusion

Model 1 is the best model due to the cross validation results, the AIC results, and the significance of the coefficint values.








