---
title: "Untitled"
author: "Kiera Matiska"
date: "2/5/2022"
output:
  html_document:
    code_folding: hide
---

## Overview:

This code runs 2 Non-Linear Least Squares Regression analysis on a model considering that weight of a species is a function of its snout to vent length. The first regression is run on the whole data set with all the species of lizards includes, whereas the second regression looks only at male Western Whiptail Lizards. Then, both models will be compared to the later regression and plotted on a graph.

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

# attach packages
library(tidyverse)
library(kableExtra)
library(equatiomatic)
library(purrr)
```

## Read in data

```{r}
lizards <- read_csv("lizard.csv") %>% # read in csv
  mutate(sex = case_when( # expand sex variable names
    sex == "M" ~ "male",
    sex == "F" ~ "female"
  ))
```

## Start nonlinear Least Square Model

W = a*SVL^b

- W = weight
- SVL = snout to vent length
- a and b are the parameters needed to be fit

### Run aa Non-linear Least Squares Regression

```{r}
# add some log columns of the data: length and width
lizards_log <- lizards %>% 
  mutate(log_length = log(SV_length)) %>% 
  mutate(log_weight = log(weight)) %>% 
  drop_na() # drop na values

lizards_model_guess <- lm(log_weight ~ log_length, data = lizards_log) # create a guess model using log values

conceptual_model <- function(a, SV_length, b) { # define the conceptual model
  W = a * SV_length ^ b
  return(W)
}

new_model <- function (a, SV_length, b) { # define the linear equation model
  ln(W) = b * ln(a) + b * ln(SV_length)
  return(W)
}

guess_coef <- coefficients(lizards_model_guess) # find the coefficients


# run a nls test
nls_lizards <- nls(weight ~ conceptual_model(a, SV_length, b),
                   data = lizards, # use this data
                   start = list(a = 2.718^(guess_coef[1]/guess_coef[2]), # calculate coefficient a
                                b = guess_coef[2]), # coefficient b already calculated
                   trace = TRUE)

summary(nls_lizards)
```

### Model Visualization

```{r}
# create a scatter plot of the data
ggplot(data = lizards,# define data
       aes(x = SV_length, y = weight)) + 
  geom_point(aes(color = sex), # define scatter plot and color points by sex
             size = 0.7) + # change point size
  labs(title = "Nonlinear Least Squares Model Comparing Lizard\nSnout to Vent Length to Weight by Sex", #add title
        x = "Snout to Vent Length (mm)", # add x-axis label
        y = "Weight (g)") + # add y-axis label
  theme(plot.title = element_text(hjust = 0.5), # center title
        plot.background = element_rect(fill = "aliceblue")) # change plot background color
```

### Model Western Whiptail Lizard

```{r}
# CNTI is Western Whiptail
# create a CNTI subset
lizards_cnti <- lizards_log %>% 
  filter(spp == "CNTI", sex == "male") %>% # filter for the species and male
  mutate(spp = case_when( # change CNTI to the common name of lizard
    spp == "CNTI" ~ "Western Whiptail"
  ))

lizards_model_guess1 <- lm(log_weight ~ log_length, data = lizards_cnti) # create a guess model using log values

conceptual_model <- function(a, SV_length, b) { # define the conceptual model
  W = a * SV_length ^ b
  return(W)
}

new_model <- function (a, SV_length, b) { # define the linear equation model
  ln(W) = b * ln(a) + b * ln(SV_length)
  print(W)
}

guess_coef1 <- coefficients(lizards_model_guess) # find the coefficients
guess_coef1

# run a nls 
nls_cnti <- nls(weight ~ conceptual_model(a, SV_length, b),
                   data = lizards_cnti, # use this data
                   start = list(a = 2.718^(guess_coef[1]/guess_coef[2]), # calculate coefficient a
                                b = guess_coef[2]), # coefficient b already calculated
                   trace = TRUE)
```

### Data Visualization

```{r}
ggplot(data = lizards_cnti,
       aes(x = SV_length, y = weight)) +
  geom_point(color = "darkolivegreen4") +
  labs(title = "Western Whiptail Lizard Snout to Vent Length vs. Weight",
       x = "Snout to Vent Length (mm)",
       y = "Weight (g)") +
  theme(plot.title = element_text(hjust = 0.5),
        plot.background = element_rect(fill = "aliceblue"))
```

### Find RMSE for both models

```{r}
RMSE_sum_df <- data.frame(matrix(c("1",
                                    rmse(lizard$weight, predict(nls_df)),
                                    "2",
                                    rmse(lizards_cnti$weight, predict(nls_df2))),
                                    2,2,byrow=T)) %>%
  set_colnames(c("model","rmse"))
```

Coefficient a is smaller in general model and larger in the species specific model. Vice Versa for coefficient b. Risidual sum of squares for general is 15467 and 807.7 for the other. General took 7 iterations and species specific took 8.
